### 2015-10-04 article.tex
Structure plus first version of the abstract

\documentclass{article}

\usepackage{csquotes}

\begin{document}

\title{Functional vs. Object Oriented Programming}
\author{Lorenzo Corneo and Nikolaos-Ektoras Anestos \\
	\{corneo, anestos\}@kth.se}
\date{\today}
\maketitle

\begin{abstract}
Human resources and time consumption are very important factors in software development. Both developers and managers try to maximise their productivity, in order to increase as much as possible the profits or gain market advantage. We assume that choosing an appropriate programming paradigm is a critical factor that leads to saving of these resources.

The main object of investigation in this project are the lines of code needed to develop two different implementations of the same program and the level of reusability of existing code.

Functional programming has some fundamental benefits that increase programmer productivity over that when OOP languages. The purpose of this project is to demonstrate some of the benefits of functional programming over OOP, in order to persuade developers to switch to functional programming.
\end{abstract}

\section{Introduction}

\section{Comparison criteria}

\section{Experimentation results}

\section{Evaluation}

\section{Acknowledgement}

\section{Conclusion}

\begin{thebibliography}{9}
\bibitem{flink}
Apache Flink. 2015. Apache Flink: Scalable Batch and Stream Data Processing. [ONLINE] Available at: http://flink.apache.org. [Accessed 16 September 15].
\bibitem{rx}
ReactiveX. 2015. ReactiveX. [ONLINE] Available at: http://reactivex.io/. [Accessed 16 September 15].
\bibitem{fpoop}
R. Harrison, \enquote*{Comparing programming paradigms: an evaluation of functional and object-oriented programs}, Jul. 1996 [Online]. Available: http://ieeexplore.ieee.org/stamp/stamp.jsp?tp=\&arnumber=511273
\bibitem{githubapi}
GitHub Developer. 2015. Statistics | GitHub API. [ONLINE] Available at: https://developer.github.com/v3/repos/statistics/. [Accessed 16 September 15].
\bibitem{benkio}
https://github.com/benkio/ITC-Slides
\end{thebibliography}

\end{document}